---
title: "Wage Datasets Analysis Code File"
author: "Radiah Khan"
format: html
editor: visual
---

## Loading Packages

```{r}
library(openintro)
library(tidyverse)
library(ggplot2)
library(arsenal)
```

## The top university majors for men and women

```{r}
view(jobs_gender)
```

## Top university majors for women

```{r}
top_majors_of_jobs_female <- jobs_gender |>
  select(major_category, workers_female) |>
  unique() |>
  group_by(major_category) |>
  summarize(count = n()) |>
  arrange(desc(count))
top_majors_of_jobs_female
```

## Distribution of Job Majors for Women (Plot 2- Slide 3)

```{r}
ggplot(data = top_majors_of_jobs_female,
       aes(x = major_category,
           y = count)) +
  geom_col(fill = "Blue") +
  labs(title = "Distribution of female job holders by college majors",
       x = "Major Category",
       y = "Count") +
 
  coord_flip()
```

## Top university majors for Men

```{r}
top_majors_of_jobs_male <- jobs_gender |>
  select(major_category, workers_male) |>
  unique() |>
  group_by(major_category) |>
  summarize(count = n()) |>
  arrange(desc(count))
top_majors_of_jobs_male
```

## Distribution of Job Majors for Men (Plot 1 - Slide 2)

```{r}
ggplot(data = top_majors_of_jobs_male,
       aes(x = major_category,
           y = count)) +
  geom_col(fill = "Red") +
  labs(title = "Distribution of male job holders by college majors",
       x = "Major Category",
       y = "Count") +
 
  coord_flip()
```

## Comparing two data frame

```{r}
dat1 <- top_majors_of_jobs_female 
dat2 <- top_majors_of_jobs_male
summary(comparedf(dat1,dat2))
print(summary(comparedf(dat1,dat2)))
```

## Combining two data frames

```{r}
combined_inner <- merge(dat1, dat2, by = "major_category")
combined_inner
```

## Renaming columns

```{r}

colnames(combined_inner)[colnames(combined_inner) == "count.x"] <- "Female Count"
```

```{r}

colnames(combined_inner)[colnames(combined_inner) == "count.y"] <- "Male Count"
```

## Tidy Data

```{r}
combined_inner_pivoted <- combined_inner %>%
  pivot_longer(cols = ends_with("Count"),
               names_to = "count_type",
               values_to = "value")
```

## Plotting the combined pivoted data (Plot 3 - Slide 5)

```{r}

ggplot(data = combined_inner_pivoted,
       aes(x = major_category,
           y = value)) +
  facet_grid(count_type ~ .)+ 
  geom_col(width = 0.5, fill = "red")+
  labs(title = "Comparison of female and male participation in different job fields based on their majors",
       x = "Major Category",
       y = "Total Count")+
  scale_x_discrete()+
  coord_flip() 
        
```

```{r}

jobs_by_gender <- jobs_gender |>
  mutate(percent_male = (workers_male/total_workers)*100)

```

## Data Wrangling

```{r}

jobs_by_gender_selected <- jobs_by_gender |> 
  select(year, major_category, total_earnings_male, total_earnings_female, percent_male, percent_female, total_earnings_female, total_earnings_male)
  
```

## Tidy Data

```{r}

jobs_by_gender_by_pivoted <- jobs_by_gender_selected |>
pivot_longer(cols = c("total_earnings_male", "total_earnings_female", "percent_male", "percent_female"),
             names_to = "category",
             values_to = "total_value")

```

# Data Wrangling

```{r}

jobs_by_gender_cleaned_total <- jobs_by_gender_by_pivoted |>
  filter(category %in% c("total_earnings_male", "total_earnings_female"))

```

```{r}
library(ggplot2)
library(ggthemes)  
library(RColorBrewer)
```

## Converting year to a factor to set the continuous to discrete scale (Plot 4- Slide 6)

```{r}
jobs_by_gender_cleaned_total$year <- as.factor(jobs_by_gender_cleaned_total$year)

```

```{r}
graph1 <- ggplot(data = jobs_by_gender_cleaned_total, 
                  aes(x = major_category, y = total_value, color = year)) + 
  geom_point(size = 3 ) + 
  facet_wrap(~category) + 
  labs(title = "Comparison of Earnings of Females and Males in Different Majors",
       x = "Major Category",
       y = "Total Earnings") +
  coord_flip() + 
    scale_color_manual(values = c("black", "red", "blue", "green"))

print(graph1)

```

## Determining age group (Plot 5- Slide 8)

```{r }
earnings_female |>
  select(group) |>
  unique()
print(earnings_female)
```

## Renaming age groups by combining them

```{r}

plot_data <- combined |> 
  mutate(broad_group = case_when(
    group %in% c("16-19 years", "20-24 years", "25-34 years") ~ "Age_group_A_16-34",
    group %in% c("35-44 years", "45-54 years", "55-64 years") ~ "Age_group_B_35-64",
    TRUE ~ NA
  )) |> 
  filter(!is.na(broad_group))
  plot_data
```

## Plotting different age group of women's data (Plot 6- Slide 9)

```{r}
plot_data |>
group_by(year, broad_group) |>
ggplot(aes(x = year, y = percent, color = broad_group)) +
geom_line() +
  labs(title = "Earnings of Women Across Different Age Groups Over the Years",
       x = "Year",
       y = "Percentage")
scale_color_brewer(palette = "Dark2")
plot_data
  
```

```{r}

plot_data_filtered_1 <- plot_data |>
  select(year, part_time_female, part_time_male)
```

## (Plot 7- Slide 10)

```{r}

ggplot(plot_data_filtered_1, aes(x = year, y = part_time_female)) +
  geom_line(color = "black", size = 1) +  
  geom_point(color = "red", size = 3) +  
  labs(title = "Percentage of Part-Time Female Jobs Over the Years",
       x = "Year",
       y = "Percentage") +
  theme_minimal()
```

## (Plot 8- Slide 11)

```{r}

ggplot(plot_data_filtered_1, aes(x = year, y = part_time_male)) +
  geom_line(color = "black", size = 1) +  
  geom_point(color = "red", size = 3) +  
  labs(title = "Percentage of Part-Time Male Jobs Over the Years",
       x = "Year",
       y = "Percentage") +
  theme_minimal()
         
```

##Data Wrangling

```{r}
plot_data_filtered_2 <- plot_data |>
  select(year, full_time_female, full_time_male)
```

## Plotting (Plot 9- Slide 12)

```{r}

ggplot(plot_data_filtered_2, aes(x = year, y = full_time_female)) +
  geom_line(color = "black", size = 1) +  
  geom_point(color = "blue", size = 3) +  
  labs(title = "Percentage of Full-Time Female Jobs Over the Years",
       x = "Year",
       y = "Percentage") +
  theme_minimal()
```

## (Plot 10- Slide 13)

```{r}

ggplot(plot_data_filtered_2, aes(x = year, y = full_time_male)) +
  geom_line(color = "black", size = 1) +  
  geom_point(color = "blue", size = 3) +  
  labs(title = "Percentage of Full-Time Male Jobs Over the Years",
       x = "Year",
       y = "Percentage") +
  theme_minimal()
```

## Black_White_Wage_Gap

```{r}

view(black_white_wage_gap)
```

```{r}

black_white_women_wage_gap <- black_white_wage_gap |>
  select(year, black_women_average, white_women_average)
```

```{r}

black_white_men_wage_gap <- black_white_wage_gap |>
  select(year, black_men_average, white_men_average)
```

## Multiple line in one plot (Plot 11- Slide 14)

```{r}
p1 <- ggplot(data = black_white_women_wage_gap, aes(x = year)) +
  geom_line(aes(y = black_women_average, color = "Average Wage Of Black Women")) +
  geom_point(aes(y = black_women_average, color = "Average Wage Of Black Women")) +
  geom_line(aes(y = white_women_average, color = "Average Wage of White Women")) +
  geom_point(aes(y = white_women_average, color = "Average Wage of White Women")) +
  theme_classic() +
  scale_colour_manual(name = "Variables", 
                      values = c("Average Wage Of Black Women" = "blue", 
                                 "Average Wage of White Women" = "red")) +
  labs(title = "Comparison of Wages Over The Years for Black and White Women",
       x = "Years",
       y = "Average Wage")

p1

```

## (Plot 12- Slide 15)

```{r}

p2 <- ggplot(data = black_white_men_wage_gap, aes(x = year)) +
  geom_line(aes(y = black_men_average, color = "Average Wage Of Black Men")) +
  geom_point(aes(y = black_men_average, color = "Average Wage Of Black Men")) +
  geom_line(aes(y = white_men_average, color = "Average Wage of White Men")) +
  geom_point(aes(y = white_men_average, color = "Average Wage of White Men")) +
  theme_classic() +
  scale_colour_manual(name = "Variables", 
                      values = c("Average Wage Of Black Men" = "blue", 
                                 "Average Wage of White Men" = "red")) +
  labs(
       x = "Years",
       y = "Average Wage")

p2


```
